---
title: "QC TME Pilka"
author: "Mohmed Abdalfttah"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    toc: yes
    toc_float: yes
    number_sections: yes
    df_print: paged
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message = FALSE, warning = FALSE)
options(width = 1200)
```

```{r}
Analysis = function(obj, res, dim){
  obj <-  SCTransform(obj, vst.flavor = "v2", verbose = TRUE, vars.to.regress = c("percent.mt"))
  #obj <- NormalizeData(obj)
  #obj <- FindVariableFeatures(obj, selection.method = "vst", nfeatures = 3000)
  #obj <- ScaleData(obj)
  obj <- RunPCA(obj, features = VariableFeatures(object = obj))
  print(ElbowPlot(object = obj, ndims = 50))
  obj = RunUMAP(object = obj, dims = dim, verbose = F)
  obj = FindNeighbors(object = obj, dims =  dim, verbose = F)
  obj = FindClusters(object = obj, resolution = res, verbose = F)
}

```

## Libraries

```{r}
library(Seurat)
library(cowplot)
library(tidyverse)
library(glue)
library(here)
```

## Setting parameters
Loading necessary paths and parameters
```{r}
source(here::here("misc/paths.R"))
```

```{r}
se_obj <- "{sc}/TME/pelka_stromal_raw.rds" %>%
  here::here() %>%
  glue::glue() %>%
  readRDS(.)
```

```{r}
se_obj <- subset(se_obj, MMRStatus %in% c("MMRp"))
```

# Quality Control 

```{r, fig.width=15}
se_obj$project.name <- "SeuratProject"
# The [[ operator can add columns to object metadata. This is a great place to stash QC stats
se_obj[["percent.mt"]] <- PercentageFeatureSet(se_obj, pattern = "^MT-")
se_obj[["percent.rb"]] <- PercentageFeatureSet(se_obj, pattern = "^RP[SL]")

```

```{r}
# Visualize QC metrics as a violin plot
VlnPlot(se_obj, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, group.by = "project.name")
VlnPlot(se_obj, features = c("nFeature_RNA", "nCount_RNA", "percent.rb"), ncol = 3, group.by = "project.name")

plot1 <- FeatureScatter(se_obj, feature1 = "nCount_RNA", feature2 = "percent.mt", group.by = "project.name")
plot2 <- FeatureScatter(se_obj, feature1 = "nCount_RNA", feature2 = "nFeature_RNA", group.by = "project.name")
plot1 + plot2

```


```{r, fig.width=20, fig.height=10}
# Visualize QC metrics as a violin plot
VlnPlot(se_obj, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, group.by = "orig.ident")
VlnPlot(se_obj, features = c("nFeature_RNA", "nCount_RNA", "percent.rb"), ncol = 3, group.by = "orig.ident")

plot1 <- FeatureScatter(se_obj, feature1 = "nCount_RNA", feature2 = "percent.mt", group.by = "orig.ident")
plot2 <- FeatureScatter(se_obj, feature1 = "nCount_RNA", feature2 = "nFeature_RNA", group.by = "orig.ident")
plot1 + plot2

```

# Filter Mito and ribosomal genes
```{r}
se_obj <- se_obj[!grepl("^MT-", rownames(se_obj)), ]
se_obj <- se_obj[ ! grepl('^RP[SL]', rownames(se_obj)), ]

```

# Add some metadata

```{r}
se_obj$Survival <- 'Alive'
se_obj$Metastasis <- 'No'
```

```{r}
gc()
```

```{r}
se_obj <- Analysis(obj = se_obj, res = 0.2, dim = 1:30)
```

```{r}
gc()
```

SpecimenType: Normal and Tumor

TumorStage: In cancer staging, "T4" refers to the size and/or extent of the primary tumor. The "T" in T4 stands for "tumor," and the "4" indicates a relatively advanced stage.

Grade: high >> T4, Low >> not T4

LymphNodeStatus: N+ (Positive Lymph Node Status): This indicates that cancer cells have spread to nearby lymph nodes. N- (Negative Lymph Node Status): This means that no cancer cells were found in the sampled lymph nodes.


```{r, fig.width= 20, fig.height=10}
colnames(se_obj@meta.data)
DimPlot(se_obj, group.by = c("SpecimenType", "Hospital", "ProcessingMethod", "Sex", "Side", "Grade", "TumorStage", "LymphNodeStatus", "MMRStatus", "MMRMLH1", "DiseaseOntology"), cols = ggsci::pal_igv()(50),raster = F)
```
```{r, fig.width=20, fig.height=10}
DimPlot(se_obj, group.by = c("Patient"), ncol = 1, cols = c(ggsci::pal_igv()(50), ggsci::pal_uchicago(palette = "light")(6), ggsci::pal_uchicago(palette = "dark")(6)),raster = F)
```


```{r}
sessionInfo()
```

```{r}
"{qc_sc}/{robj_dir}/MSS_Pellka.rds" %>%
  here::here() %>%
  glue::glue() %>%
  saveRDS(object = se_obj)


```

